---
title: "Statistical assignment 4"
author: "Sunny Chou"
date: "01/03/2020"
output: github_document
---

```{r setup, include=FALSE}
# Please note these options.
# This tells R Markdown that we want to show code in the output document.
knitr::opts_chunk$set(echo = TRUE)
# Switching off messages in the output document. 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
# Switching on caching to make things faster (don't commit cache files on Github).
knitr::opts_chunk$set(cache = TRUE)
```

In this assignment you will need to reproduce 5 ggplot graphs. I supply graphs as images; you need to write the ggplot2 code to reproduce them and knit and submit a Markdown document with the reproduced graphs (as well as your .Rmd file).

First we will need to open and recode the data. I supply the code for this; you only need to change the file paths.

    ```{r}
library(tidyverse)
library(ggplot2)
    
getwd()

Data8 <- read_tsv("C:/Users/user/Desktop/data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")
Data8 <- Data8 %>%
        select(pidp, h_age_dv, h_payn_dv, h_gor_dv)
Stable <- read_tsv("C:/Users/user/Desktop/data/UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab")
Stable <- Stable %>%
        select(pidp, sex_dv, ukborn, plbornc)
Data <- Data8 %>% left_join(Stable, "pidp")
rm(Data8, Stable)
Data <- Data %>%
        mutate(sex_dv = ifelse(sex_dv == 1, "male",
                           ifelse(sex_dv == 2, "female", NA))) %>%
        mutate(h_payn_dv = ifelse(h_payn_dv < 0, NA, h_payn_dv)) %>%
        mutate(h_gor_dv = recode(h_gor_dv,
                         `-9` = NA_character_,
                         `1` = "North East",
                         `2` = "North West",
                         `3` = "Yorkshire",
                         `4` = "East Midlands",
                         `5` = "West Midlands",
                         `6` = "East of England",
                         `7` = "London",
                         `8` = "South East",
                         `9` = "South West",
                         `10` = "Wales",
                         `11` = "Scotland",
                         `12` = "Northern Ireland")) %>%
        mutate(placeBorn = case_when(
                ukborn  == -9 ~ NA_character_,
                ukborn < 5 ~ "UK",
                plbornc == 5 ~ "Ireland",
                plbornc == 18 ~ "India",
                plbornc == 19 ~ "Pakistan",
                plbornc == 20 ~ "Bangladesh",
                plbornc == 10 ~ "Poland",
                plbornc == 27 ~ "Jamaica",
                plbornc == 24 ~ "Nigeria",
                TRUE ~ "other"))

    ```

Reproduce the following graphs as close as you can. For each graph, write two sentences (not more!) describing its main message.  

1. Univariate distribution (20 points).

    ```{r}
    
ggplot(Data, aes(x = h_payn_dv)) +
  geom_freqpoly() +
  xlab("Net Monthly Pay") + ylab('Number of Respondents')
 
    ```

Most the the respondents, around 2500 of them, has the net monthly pay between 1000 to 1500. The trend steeply rises to that point and steeply 
declines until a surge of a small amount of respondents with income around 5500.



2. Line chart (20 points). The lines show the non-parametric association between age and monthly earnings for men and women.

    ```{r}
library(gcookbook)
plot2<- ggplot(Data, aes(x = h_age_dv, y = h_payn_dv , linetype=sex_dv)) + geom_smooth()+ xlim(16,65) + xlab("Age") + ylab('Monthly Earnings')
plot2

    ```
Both male and female monthly earnings exhibit a similiar trend, steepy rising towards a more moderate rise, and falls gradually around 45 to 50 respectively. However, female wages rise at a much slower rate after 25, and were capped at 1500.



3. Faceted bar chart (20 points).

    ```{r}
    
    Plot3data1 <- Data %>% group_by(placeBorn,sex_dv) 
    Plot3data2 <- Plot3data1 %>% summarise(MedianIncome = median(h_payn_dv, na.rm = TRUE))
    plot3data<-na.omit(Plot3data2)
    
    plot3<-ggplot(plot3data, aes(x= sex_dv, y = MedianIncome)) + geom_bar(stat = "identity") + ylim(0,2000)
    plot3<-plot3 + facet_wrap(.~placeBorn) + ylab('Median Monthly Net Pay') + xlab('Sex')
    plot3
    

    
    ```

Median monthly net pay for male has been higher than female across all countries of origin. More developed countries of Birth have shown greater disparity between male and female pay, with male higher.


4. Heat map (20 points).

    ```{r}
    
     Plot4data <- Data %>% group_by(placeBorn,h_gor_dv) %>% summarise(MeanAge = mean(h_age_dv, na.rm = TRUE))
     Plot4data <- na.omit(Plot4data)
     plot4 <- ggplot(Plot4data, aes(x = h_gor_dv, y = placeBorn, fill = MeanAge))  + xlab('Region') + ylab('Countries of Birth')
     plot4 <- plot4 + geom_tile() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
     plot4
    
    ```
Mean age in the UK looks even dispersed across regions and countries of Birth, with Nigeria the youngest.

5. Population pyramid (20 points).

    ```{r}
    
 plot5data <- Data %>% group_by(h_age_dv, sex = sex_dv) %>% count(age = h_age_dv) 

 plot5data <- na.omit(plot5data)
  
 plot5 <- ggplot(data = plot5data, mapping = aes(x = h_age_dv, y = ifelse(test = sex == "male", yes = -n, no = n), 
                     fill = sex)) + geom_col() + coord_flip() + 
     scale_y_continuous(labels = abs, limits = max(plot5data$n) * c(-1,1))  + xlab('Age')+ ylab('n')
      
 plot5
    
    
    ```


Female distribution greater than male between 25 to 50, also between 67.5 to 75. Otherwise similarly distributed.

